Teaching Debugging Collaboratively---Midway Report

### 2020-05-10 paper/midway.tex
some background

\documentclass[12pt]{article}
\usepackage{multicol}
\usepackage{cite}

@@

>>>>

\documentclass[12pt]{article}
\usepackage{multicol}
\usepackage{cite}
\usepackage{listings}

\lstset{basicstyle=\ttfamily}

@@

\subsection{Overview of RR}

Time traveling debugging---stepping backwards and forwards through a
program's execution---is supported naively by GDB through the use of
commands such as \lstinline{reverse-continue} and
\lstinline{reverse-stepi}.\cite{gdbman} What RR brings to the table is
the ability to deterministically record and replay the execution of
even very complex programs.  This allows single users to capture and
debug events that may not happen every time a program runs and revisit
debugging later.  Essentially for my planned collaborative debugger,
execution can be replayed and debugged even \textit{on a different
  computer} than the computer that recorded the original execution.
In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
that rely on capturing the entire state of of a virtual machine to
replay execution, RR records and replays faster, produces far smaller
files, and doesn't force execution inside of a
VM.\cite{DBLP:journals/corr/OCallahanJFHNP17}

\subsection{Limitations of RR}

RR comes with two major limitations: it is only compatible with the
Linux kernel, and it's deterministic recording and replay relies on a
feature that is only found on modern \textit{Intel} x86 CPUs.  For my
purposes the first limitation is fairly irrelevant--I expect students
to be connecting to a central server in order to execute and replay
code, and most systems classes are taught in Linux.  The second
limitation poses more of a problem for instruction.  Because recording
and replay is limited to x86 CPUs, classes cannot use a debugger based
on RR to teach ARM assembly, which is becoming increasingly popular.
Though this isn't a limitation that solutions like the aforementioned
full-vm recorder PANDA face, RR still maintains the advantage of being
a debugger first and foremost.  Students can learn RR, GDB, and
debugging techniques simultaneously.  Platforms that require entire VM
replay come with far greater performance and instructional overhead.
Because of RR's ease of use and tight integration with GDB, I feel it
is the best choice of platform to base a collaborative debugger on.
